---
title: "practice_exam_2"
author: "Carter Martindale"
date: "10/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fivethirtyeight)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(rsample)
library(rstanarm)
```

## Question 1

```{r q1}
US_births_2000_2014 %>% 
  group_by(year, month) %>% 
  summarise(mean_births = mean(births), .groups = "drop") %>% 
  ggplot(aes(x = month, y = mean_births)) +
  geom_line() +
  facet_wrap(~year) +
  scale_x_continuous(breaks = c(1, 6, 12),
                     labels = c("Jan.", "June",
                                "Dec.")) +
  labs(title = "Number of US Births by Month",
       x = "Month",
       y = "Mean # of Births") +
  theme_minimal() +
  theme(panel.spacing = unit(0.8, "lines"))
```

## Question 2

```{r q2}

random_sample <- sample_n(US_births_2000_2014, size = 500,
                          replace = TRUE)

nonrandom_sample <- US_births_2000_2014 %>% 
  filter(month == 9)

avg_births_plain <- US_births_2000_2014 %>% 
  summarise(avg_births = mean(births))

avg_births_random <- random_sample %>% 
  summarise(avg_births = mean(births))

avg_births_nonrandom <- nonrandom_sample %>% 
  summarise(avg_births = mean(births))

```
## Question 3
Then, run this function on your random_sample and nonrandom_sample objects separately and store the results into objects called boot_random and boot_nonrandom. Finally, using the quantile function on each of these objects, create 95% confidence intervals for your estimates and store them in objects called boot_ci_rand and boot_ci_nonrand. Do either of your confidence intervals contain the true value? Which one?
```{r q3}
bootstrap_estimate <- function(x) {
  x %>% 
    bootstraps(times = 1000) %>% 
    mutate(boot = map(splits, ~ analysis(.))) %>% 
    mutate(birth_boot = map(boot, ~ pull(., births))) %>% 
  mutate(avg_births = map_dbl(birth_boot, ~ mean(.)))
}

boot_random <- bootstrap_estimate(random_sample)
boot_nonrandom <- bootstrap_estimate(nonrandom_sample)

boot_ci_rand <- quantile(boot_random$avg_births, probs = c(0.025, 0.975))
boot_ci_nonrand <- quantile(boot_nonrandom$avg_births, probs = c(0.025, 0.975))
```

## Question 4
```{r q4}
boot_random_filtered <- boot_random %>% 
  select(avg_births)
boot_nonrandom_filtered <- boot_nonrandom %>% 
  select(avg_births)

boot_combined <- bind_rows(boot_random_filtered,
                           boot_nonrandom_filtered,
          .id = "sample_type")

ggplot(boot_combined, aes(x = avg_births,
                          y = after_stat(count/sum(count)),
                          fill = sample_type)) +
  geom_histogram(bins = 50) +
  geom_vline(xintercept = 11350.07, lty = "dotted") +
  scale_fill_manual(name = "Sample Type",
                    values = c("salmon", "dodgerblue"),  
                      labels = c("Random", "Nonrandom")) +
  theme_minimal() +
  labs(title = "Posterior Distribution of Average Births",
       subtitle = "Random Samples are More Reliable",
       x = "Average Births",
       y = "Probability")
```
## Question 5
```{r q5}

stan_random <- stan_glm(data = boot_random,
         avg_births ~ 1,
         refresh = 0)

stan_nonrandom <- stan_glm(data = boot_nonrandom,
                           avg_births ~ 1,
                           refresh = 0)

rand_ci <- posterior_interval(stan_random, prob = 0.95)
nonrand_ci <- posterior_interval(stan_nonrandom, prob = 0.95)
stan_random
```

